---
title: <center><font size="6"><b>Calibration</b></font></center>
subtitle: <center><font size="4"><b>Long-billed hermit song amplitude</b></font></center>
author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a> & Melanie Talavera</font></center> 
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

#### Load packages
```{r Load packages, eval = TRUE, message=FALSE, warning = FALSE}

## vector with package names
x <- c( "pbapply", "parallel", "ggplot2", "warbleR", "Rraven", "readxl", "Sim.DiffProc", "png", "baRulho", "viridis", "sysfonts")

aa <- lapply(x, function(y) {
  
  # check if installed, if not then install 
  if (!y %in% installed.packages()[,"Package"]) 
    install.packages(y) 

  # load package
  try(require(y, character.only = T), silent = T)
})

knitr::opts_knit$set(root.dir = normalizePath(".."), warnings = FALSE, message = FALSE)
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)

font_add(family = "Calibri light", regular = "./data/calibril.ttf")
```

```{r, parameters and functions, eval = TRUE}

calibration.path <- "./data/raw/calibration"

# parameters for calibration
f = 48
reps <- 5
gap.dur <- 1

# for ggplots
pd <- position_dodge(0.1)

lseq <- function(from=1, to=100000, length.out=6) {
  # logarithmic spaced sequence
  # blatantly stolen from library("emdbook"), because need only this
  exp(seq(log(from), log(to), length.out = length.out))
}

decrease <- (1 / lseq(length.out = 11))[-11]

wav_dur <- wavdur

detec_saturation <- function(X, parallel = 1, bit = 16, max_amplitude = NULL, path = NULL){
  
  if (is.null(max_amplitude))
  max_amplitude <- max((2 ^ bit) / 2) - 1
  
  out <- pblapply(1:nrow(X), cl = parallel, function(x){
    
    wv <- read_wave(X, index = x, path = path)
  
    out_df <- data.frame(sound.files = X$sound.files[x], selec = X$selec[x], prop.saturated = sum(wv@left == max_amplitude) / length(wv@left))
    
    return(out_df) 
})

 sat_df <- do.call(rbind, out)
 
 
 if (any(sat_df$prop.saturated > 0.2))
 cat(crayon::magenta(paste(sum(sat_df$prop.saturated > 0.2), "selections look saturated (i.e. > 20% of amplitude samples reached the highest value)"))) 

  if (any(all(sat_df$prop.saturated < 0.2) & any(sat_df$prop.saturated > 0.05)))
 cat(crayon::magenta(paste(sum(sat_df$prop.saturated > 0.05), "selections with some degree of saturation (i.e. < 20% but  > 5% of amplitude samples reached the highest value)"))) 

  if (all(sat_df$prop.saturated < 0.05))
 cat(crayon::silver("no saturation detected (all selections < 5% of amplitude samples reached the highest value)")) 

 
  return(sat_df)

}


```

# Measure frequency on real data
```{r measure frequency on real data, eval = FALSE}

ad.test <- auto_detec(threshold = 25, bp = c(1, 9), ssmooth = 300, path = "./data/processed/recording_cuts", flist = c("397_SUR_02-Feb-2021_09.21_10.32_AAE-6.wav", "416_SUR_08-Feb-2021_07.36_10.47_A3C-4.wav", "417_SUR_27-Jan-2021_07.27_10.14_AED-1.wav", "438_SJA_12-Feb-2021_08.21_09.19_AAE-2.wav", "432_SJA_16-Feb-2021_08.21_11.01_ABB-3.wav", "437_SJA_10-Feb-2021_07.30_10.46_A1A-3.wav", "444_CCL_27_Feb-2021_08.05_08.32_AC8-1.wav", "446_CCL_27-Feb-2021_07.17_07.48_A3C-1.wav", "448_CCL_26-Feb-2021_00.00_00.00_AC8-2.wav"), thinning = 0.05, hold.time = 0.1)

sa <- spectro_analysis(ad.test,  path = "./data/processed/recording_cuts", bp = c(1, 9), ssmooth = 300)

write.csv(sa, file.path(calibration.path, "spectro_analysis_on_subset_for_peak_frequency_for_calibration.csv"), row.names = FALSE)

```

# Amplitude calibration
```{r creating playback for soundmeter calibration, eval = FALSE}

source('~/Dropbox/R_package_testing/baRulho/baRulho/R/master_sound_file.R')
source('~/Dropbox/R_package_testing/warbleR/warbleR/R/sim_songs.R')
source('~/Dropbox/R_package_testing/warbleR/warbleR/R/fade_env_wrblr_int.R')

# remove file created in next line (it cannot be overwritten)
unlink(file.path(calibration.path, "tones_for_amplitude_test_soundmeter.wav"))

# simulate sounds
sm_sng_sm <- sim_songs(freq = 6.5, n = length(decrease), harms = 1, diff.fun = "pure.tone", bgn = 0.0001, fin = 0.2, fout = 0.2, shape = "cos", selec.table = TRUE, samp.rate = f, file.name = "tones_for_amplitude_test_soundmeter", path = calibration.path, durs = 5, gaps = gap.dur)

# normalize
sm_sng_sm$wave <- normalize(sm_sng_sm$wave, unit = "16")

# reduce amplitude 
for (i in 1:nrow(sm_sng_sm$selec.table))
sm_sng_sm$wave@left[(sm_sng_sm$selec.table$start[i]*f*1000):(sm_sng_sm$selec.table$end[i]*f*1000)] <- sm_sng_sm$wave@left[(sm_sng_sm$selec.table$start[i]*f*1000):(sm_sng_sm$selec.table$end[i]*f*1000)] * decrease[i]

# seewave::spectro(sm_sng_sm$wave, scale = FALSE, osc = TRUE, heights = c(1, 1), grid = FALSE, flim = c(5, 8), collevels = seq(-100, 0 , 5), palette = reverse.gray.colors.1, ovlp = 0, fast.spec = TRUE)

saveRDS(sm_sng_sm, "./data/raw/calibration/tones_for_amplitude_test_soundmeter.RDS")

writeWave(sm_sng_sm$wave, "./data/raw/calibration/tones_for_amplitude_test_soundmeter.wav", extensible = FALSE)

```

```{r creating playback for audiomoth calibration, eval = FALSE}

# remove file created in next line (it cannot be overwritten)
unlink(file.path(calibration.path, "tones_for_amplitude_test_audiomoths.wav"))

# simulate sounds
sm_sng_ams <- sim_songs(freq = 6.5, n = length(decrease), harms = 1, diff.fun = "pure.tone", bgn = 0.0001, fin = 0.05, fout = 0.05, shape = "cos", selec.table = TRUE, samp.rate = f, file.name = "tones_for_amplitude_test_audiomoths", path = calibration.path)

# normalize
sm_sng_ams$wave <- normalize(sm_sng_ams$wave, unit = "16")

# reduce amplitude 
for (i in 1:nrow(sm_sng_ams$selec.table))
sm_sng_ams$wave@left[(sm_sng_ams$selec.table$start[i]*f*1000):(sm_sng_ams$selec.table$end[i]*f*1000)] <- sm_sng_ams$wave@left[(sm_sng_ams$selec.table$start[i]*f*1000):(sm_sng_ams$selec.table$end[i]*f*1000)] * decrease[i]

# seewave::spectro(sm_sng_ams$wave, scale = FALSE, osc = TRUE, heights = c(1, 1), grid = FALSE, flim = c(5, 8), collevels = seq(-100, 0 , 5), palette = reverse.gray.colors.1, ovlp = 0, fast.spec = TRUE)

saveRDS(sm_sng_ams, file.path(calibration.path, "tones_for_amplitude_test_audiomoths.RDS"))

writeWave(sm_sng_ams$wave, file.path(calibration.path, "tones_for_amplitude_test_audiomoths.wav"), extensible = FALSE)

```

```{r creating playback for soundmeter and audiomoth calibration, eval = FALSE}

sm_sng_sm <- readRDS(file.path(calibration.path, "tones_for_amplitude_test_soundmeter.RDS"))

sm_sng_ams <- readRDS(file.path(calibration.path, "tones_for_amplitude_test_audiomoths.RDS"))

sm_sng <- list()
sm_sng$wave <- pastew(sm_sng_ams$wave, sm_sng_sm$wave, output = "Wave", f = f * 1000)

writeWave(sm_sng$wave, file.path(calibration.path, "tones_for_amplitude_test_audiomoths_and_soundmeter.wav"), extensible = FALSE)

# add duration of tones for soundmeter to tones for audiomoth
# sm_sng_ams$selec.table$end <- sm_sng_ams$selec.table$end + duration(sm_sng_sm$wave)
# 
# sm_sng_ams$selec.table$start <- sm_sng_ams$selec.table$start + duration(sm_sng_sm$wave)

sm_sng$selec.table <- rbind(sm_sng_sm$selec.table, sm_sng_ams$selec.table)

# plot spectro
# seewave::spectro(clp, scale = FALSE, osc = TRUE, heights = c(1, 1), grid = FALSE, flim = c(5, 8), collevels = seq(-100, 0 , 5), palette = reverse.gray.colors.1)

## replicate audiomoths sound file
# extract selection table
st_ams <- sm_sng_ams$selec.table

st_ams2 <- st_ams[1, ]
st_ams2$end <- st_ams$end[nrow(st_ams)]

for(i in 1:reps)
st_ams2 <- rbind(st_ams2, st_ams2)
st_ams2$selec <- 1:nrow(st_ams2)


## replicate soundmeter sound file
# extract selection table
st_sm <- sm_sng$selec.table[sm_sng$selec.table$sound.files != "tones_for_amplitude_test_audiomoths.wav", ]

st_sm <- sm_sng_sm$selec.table

st_sm2 <- st_sm[1, ]
st_sm2$end <- st_sm$end[nrow(st_sm)]

for(i in 1:4)
st_sm2 <- rbind(st_sm2, st_sm2)

st2 <- rbind(st_sm2, st_ams2)

# add freq range (0.5 kHz)
st2$bottom.freq <- st2$bottom.freq - 0.3
st2$top.freq <- st2$top.freq + 0.3

st2$selec <- 1:nrow(st2)

# st2$sound.files <- "tones_for_amplitude_test_audiomoths_and_soundmeter.wav"

# make an extended selection table
synth.est <- selection_table(X = st2, extended = TRUE, pb = FALSE, confirm.extended = FALSE, path = calibration.path, by.song = "sound.files")

# create master sound file
synth.master.sf <- master_sound_file(X = synth.est, file.name = "calibration_sound_for_audiomoths_and_soundmeter", gap.duration = 1, dest.path = calibration.path, overwrite = TRUE, delay = 0, amp.marker = 1)


out <- pblapply(1:nrow(synth.master.sf), function(x){
  
  if(grepl("marker", synth.master.sf$orig.sound.file[x])){
    out <- synth.master.sf[x, ]
    out$label <- synth.master.sf$orig.sound.file[x]
    } else {
      
      if (grepl("soundmeter", synth.master.sf$orig.sound.file[x]))
       st_temp <- st_sm else st_temp <- st_ams 
      
       # make them start at 0
      st_temp$end <- st_temp$end - min(st_temp$start) 
      st_temp$start <- st_temp$start - min(st_temp$start) 
      st_temp$start <- st_temp$start + synth.master.sf$start[x]  
      st_temp$end <- st_temp$end + synth.master.sf$start[x]  
      
      st_temp$label <- round(decrease, 4)
      st_temp$orig.sound.file <- synth.master.sf$orig.sound.file[x]
      out <- st_temp
    }
  return(out)
} )

st3 <- do.call(rbind, out)

st3$bottom.freq <- st3$bottom.freq - 0.25
st3$top.freq <- st3$top.freq + 0.25

st3$selec <- 1:nrow(st3)

# st3$label <- c("start", rep(seq(1, 0.01, length.out = 10), (nrow(st3) - 2) / 10), "end")
st3$sound.files <- synth.master.sf$sound.files[1]

full_spectrograms(st3, path = calibration.path, fast.spec = TRUE, labels = "label", sxrow = 8, rows = 34, flim = c(3, 8), ovlp = 0)

write.csv(st3, file.path(calibration.path, "selection_table_audiomoths_and_soundmeter_playback.csv"), row.names = FALSE)
```

## Sound pressure level from soundmeter for synthetic tones
 
 - Synthetic sounds were created as decreasing amplitude in a log scale
 
```{r measuring sound pressure on soundmeter calibration sound, eval = TRUE}
calb <- read_excel(file.path(calibration.path, "soundmeter_calibration.xls"))

plot(1:100, calb$`Noise(dB)`[1:100], xlab = "Time (s)", ylab = "SPL (dB)", col = viridis(10)[3], pch = 20, cex = 2)

# xys <- locator(1)

dbdf <- data.frame(dB =calb$`Noise(dB)`[rep(seq(1, 2000, 6) + 8, each = 3) + 0:2])
dbdf$label <- rep(rep(round(decrease, 5), each = 3), 40)[1:nrow(dbdf)]

# select those that match noise (the rest (> 450) don't match labels)
dbdf <- dbdf[1:450,]

agg_spl_sm <- aggregate(dB ~ label, data = dbdf, mean)

agg_spl_sm$sd <- aggregate(dB ~ label, data = dbdf, sd)$dB
agg_spl_sm$sound.files <- "soundmeter"
agg_spl_sm$label <- c(0.0000, 0.0010, 0.0032, 0.0100, 0.0316, 0.1000, 0.3162, 1.0000, 0.0001, 0.0003)
```

## Mean (+/- sd) of sound pressure level from soundmeter for each synthetic tone
```{r measuring sound pressure on soundmeter calibration sound 2, eval = TRUE}
agg_spl_sm$label_cat <- factor(agg_spl_sm$label, levels = agg_spl_sm$label)

ggplot(agg_spl_sm, aes(x = label_cat, y = dB, color = sound.files)) +
    geom_errorbar(aes(ymin = dB - sd, ymax = dB + sd),width=.4, position=pd) +
    geom_line(position=pd) +
      scale_color_viridis_d(begin = 0.2, end =  0.8) +
    geom_point(position=pd, size = 2) +
  theme_classic() + labs(x = "Calibration sound decrease factor", y = "SPL (dB)")

```

```{r detecting calibration sound in audiomoth recordings, eval = FALSE}

sel_tab <- read.csv(file.path(calibration.path, "selection_table_audiomoths_and_soundmeter_playback.csv"))

frng_start <- freq_range(sel_tab[1, ], path = calibration.path)

# adjust start freq range
sel_tab$bottom.freq[1] <- frng_start$bottom.freq
sel_tab$top.freq[1] <- frng_start$top.freq

wavs <- list.files(path = calibration.path, pattern = "\\.WAV$")

found_starts <- search_templates(X = sel_tab, 
template.rows = which(sel_tab$orig.sound.file == "start_marker"), 
test.files = wavs, path = calibration.path, pb = TRUE, parallel = 1)

write.csv(found_starts, file.path(calibration.path, "starts_for_audiomoth_sync.csv"), row.names = FALSE)

```

```{r fixing delays, eval = FALSE}

found_starts <- read.csv(file.path(calibration.path, "starts_for_audiomoth_sync.csv"))

alg.tests_df <- align_test_files(X = sel_tab, Y = found_starts, path = calibration.path, by.song = TRUE, marker = "end", mar = 0.1, output = "data.frame")

exp_raven(X = alg.tests_df, path = calibration.path, sound.file.path = normalizePath(calibration.path), file.name = "audiomoth_calibration_sound_sync.txt")

fix_time <- data.frame(sound.files = c("calibration_LM_A1A.WAV", "calibration_LM_A3C.WAV", "calibration_LM_A41A.WAV", "calibration_LM_A61.WAV", "calibration_LM_AAE.WAV", "calibration_LM_ABB.WAV", "calibration_LM_AC8.WAV", "calibration_LM_AED.WAV", "calibration_M_A1A.WAV", "calibration_M_A3C.WAV", "calibration_M_A41A.WAV", "calibration_M_A61.WAV", "calibration_M_AAE.WAV", "calibration_M_AC8.WAV", "calibration_LM_A56.WAV"), 
           fix.start = c(1094.0962 - 1094.0655, 2218.2000 - 2218.1660, 3401.6597 - 3401.6116, 4658.9416 - 4658.8791, 5862.1449 - 5862.1145, 6988.0653 - 6988.0220, 8245.7745 - 8245.7395, 9359.3046 - 9359.2705, 10626.2795 - 10626.2553, 11749.9104 - 11749.8747, 12865.9948 - 12865.9454, 13977.6981 - 13977.6395, 15106.0178 - 15105.9875, 16285.0662 - 16285.0320, 4660.2555 - 4660.2103), 
           old.selec = c(472, 472, 472, 472, 472, 472, 472, 472, 472, 472, 472, 472, 472, 472, 472))

alg.tests_fix_delay <- alg.tests_df

for(i in 1:nrow(fix_time)){
      
  delay <- seq(0, fix_time$fix.start[i], length.out = fix_time$old.selec[i])
  
  delay <- c(delay, delay[2:(1 + sum(alg.tests_df$sound.files == fix_time$sound.files[i]) - fix_time$old.selec[i])] + delay[length(delay)])
  
  alg.tests_fix_delay$start[alg.tests_fix_delay$sound.files == fix_time$sound.files[i]] <- alg.tests_fix_delay$start[alg.tests_fix_delay$sound.files == fix_time$sound.files[i]] + delay  

  alg.tests_fix_delay$end[alg.tests_fix_delay$sound.files == fix_time$sound.files[i]] <- alg.tests_fix_delay$end[alg.tests_fix_delay$sound.files == fix_time$sound.files[i]] + delay  

  }

  # only for plotting and double-cheching
alg.tests_fix_delay$bottom.freq <- 3.2
alg.tests_fix_delay$top.freq <- 7.8

exp_raven(X = alg.tests_fix_delay, path = calibration.path, sound.file.path = normalizePath(calibration.path), file.name = "audiomoth_calibration_sound_sync_delay_fixed.txt")

# full_spectrograms(alg.tests_fix_delay, path = calibration.path, fast.spec = TRUE, labels = "label", sxrow = 8, rows = 60, flim = c(3, 8), ovlp = 0)

alg.tests_fix_delay$label <- sel_tab$label

alg.tests <- selection_table(X = alg.tests_fix_delay, extended = TRUE, path = calibration.path, by.song = "sound.files", mar = 0.1)

alg.tests <- rename_est_waves(alg.tests, new.sound.files = sapply(alg.tests$sound.files, function(x) strsplit(x, split = ".WAV")[[1]][1]))

saveRDS(alg.tests, file.path(calibration.path, "sel_tab_audiomoth_calibration.RDS"))

```

```{r detect saturated re-recorded calibration sounds, eval = FALSE}

alg.tests <- readRDS(file.path(calibration.path, "/sel_tab_audiomoth_calibration.RDS"))

sat_sels <- detec_saturation(X = alg.tests, parallel = 20, bit = 16)

# agg_sat <- aggregate(prop.saturated ~ sound.files + selec, data = sat_sels, mean)
# 
# agg_sat <- agg_sat[order(agg_sat$sound.files), ]
# 
# agg_sat[agg_sat$prop.saturated > 0.05, ]
# 
# agg_sat <- merge(agg_sat, saturated_df)
# 
# agg_sat[order(agg_sat$prop.saturated), ]

sum(sat_sels$prop.saturated < 0.01) / nrow(sat_sels)

unsat_sels <- alg.tests[sat_sels$prop.saturated < 0.01, ]

saveRDS(unsat_sels, file.path(calibration.path, "unsaturated_sel_tab_audiomoth_calibration.RDS"))

```

```{r calculating audiomoth calibration dBs, eval = FALSE}

alg.tests <- readRDS(file.path(calibration.path, "unsaturated_sel_tab_audiomoth_calibration.RDS"))

# alg.tests <- alg.tests[grep("marker", alg.tests$template, invert = TRUE), ]

alg.tests$spl_snr <- pbsapply(1:nrow(alg.tests), cl = 20, function(x) {

  signal <- read_wave(alg.tests, index = x, path = calibration.path)
  
  noise1 <- read_wave(alg.tests, index = x, from = alg.tests$end[alg.tests$sound.files == alg.tests$sound.files[x]][1] + 0.1, to = alg.tests$end[alg.tests$sound.files == alg.tests$sound.files[x]][1] + 0.9, path = calibration.path)
    
  signal <- ffilter(signal, f = signal@samp.rate, from = bp[1] * 1000, to = bp[2] * 1000, bandpass = TRUE, output = "Wave")
  
  noise1 <- ffilter(noise1, f = noise1@samp.rate, from = bp[1] * 1000, to = bp[2] * 1000, bandpass = TRUE, output = "Wave")
    
  # spectro(signal, flim = c(2, 10), palette = reverse.gray.colors.1, collevels = seq(-100, 0, 5))
  
  sigamp <- seewave::rms(seewave::env(signal, envt = "abs", plot = FALSE))
  noisamp <- seewave::rms(seewave::env(noise1, envt = "abs", plot = FALSE))

  signaldb <- 20 * log10(sigamp)
  noisdb <- 20 * log10(noisamp)

  spl <- 10 * log10((10 ^ (signaldb/ 10)) - (10 ^ (noisdb / 10)))

    return(spl)
  })

alg.tests$spl <- pbsapply(1:nrow(alg.tests), cl = 20, function(x) {

  signal <- read_wave(alg.tests, index = x, path = calibration.path)
  
   # signal <- ffilter(signal, f = signal@samp.rate, from = bp[1] * 1000, to = bp[2] * 1000, bandpass = TRUE, output = "Wave")
   
  sigamp <- seewave::rms(seewave::env(signal, envt = "abs", plot = FALSE))
  signaldb <- 20 * log10(sigamp)
  
    return(signaldb)
  })

# get mean and sd
agg_spl <- aggregate(spl ~ sound.files + label, data = alg.tests, meandB)
agg_spl$sd <- aggregate(spl ~ sound.files + label, data = alg.tests, sddB)$spl

# difference among H for 2 audiomoths
agg_spl$label <- as.numeric(agg_spl$label)

write.csv(agg_spl, file.path(calibration.path, "spl_by_audiomoth_calibration_sounds.csv"), row.names = FALSE)
```

## Raw amplitude values (before calibration)

Split in 3 plots to better see specific audiomoths

```{r plotting audiomoth calibration dBs, eval = TRUE}

agg_spl <- read.csv(file.path(calibration.path, "spl_by_audiomoth_calibration_sounds.csv"))

names(agg_spl)[3] <- "dB"
agg_spl_sm$label_cat <- NULL

agg_spl_all <- rbind(agg_spl_sm, agg_spl)

agg_spl_all$orig.dB <- factor(agg_spl_all$label, levels = unique(agg_spl_all$label))

ggplot(agg_spl_all[agg_spl_all$sound.files %in% c("soundmeter", unique(agg_spl_all$sound.files)[2:6]), ], aes(x = orig.dB, y = dB, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin=dB - sd, ymax = dB + sd), colour="black", width=.2, position=pd) +
  scale_color_viridis_d() +
    geom_line(position=pd) +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound")

ggplot(agg_spl_all[agg_spl_all$sound.files %in% c("soundmeter", unique(agg_spl_all$sound.files)[7:10]), ], aes(x = orig.dB, y = dB, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin=dB - sd, ymax = dB + sd), colour="black", width=.2, position=pd) +
  scale_color_viridis_d() +
    geom_line(position=pd) +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound")

ggplot(agg_spl_all[agg_spl_all$sound.files %in% c("soundmeter", unique(agg_spl_all$sound.files)[11:16]), ], aes(x = orig.dB, y = dB, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin=dB - sd, ymax = dB + sd), colour="black", width=.2, position=pd) +
  scale_color_viridis_d() +
    geom_line(position=pd) +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound")


```

```{r audiomoth pre calibration for presentation, eval = TRUE}

agg_spl_all$am <- as.character(as.numeric(as.factor(agg_spl_all$sound.files)))
agg_spl_all$am[agg_spl_all$sound.files == "soundmeter"] <- "soundmeter"
agg_spl_all$am <- factor(agg_spl_all$am, levels = c("soundmeter", 1:16))
agg_spl_all$line.size <- ifelse(agg_spl_all$am == "soundmeter", 2, 1)

ggplot(agg_spl_all[agg_spl_all$orig.dB %in% c(0.0316, 0.1, 0.3162, 1), ], aes(x = orig.dB, y = dB, group = am, color = am, size = line.size)) +
    geom_errorbar(aes(ymin= dB - sd, ymax = dB + sd, color = am),width=.2, position=pd) +
    geom_line(position=pd) +
    scale_color_viridis_d() +
    geom_point(position=pd, size=1) +
  theme_classic(base_size = 30) + labs(x = "Calibration sound SPL (dB)", y = "SPL measured (dB)", color = "") + guides(size = "none")

ggsave("./output/uncalibrated_audiomoth_spl.jpeg", width = 11, height = 7)

```



## Calibrating audiomoths using proportional adjustment

Note that this way of calibrating has high and non-randomly distributed error

```{r audiomoth calibration 3, eval = TRUE}
agg_spl <- read.csv(file.path(calibration.path, "spl_by_audiomoth_calibration_sounds.csv"))

names(agg_spl)[3] <- "dB"
agg_spl_sm$sound.files <- "soundmeter"
agg_spl_sm$label <- sort(as.numeric(unique(agg_spl$label)))

agg_spl_all <- rbind(agg_spl_sm, agg_spl)

agg_spl_all$orig.dB <- factor(agg_spl_all$label, levels = unique(agg_spl_all$label))

agg_spl_all$adj_dB <- agg_spl_all$dB

agg_spl_all <- agg_spl_all[agg_spl_all$label > 0.01, ]

for(i in unique(agg_spl_all$sound.files[agg_spl_all$sound.files != "soundmeter"]))
agg_spl_all$adj_dB[agg_spl_all$sound.files == i] <- agg_spl_all$adj_dB[agg_spl_all$sound.files == i] *
  (mean(agg_spl_all$dB[agg_spl_all$sound.files == "soundmeter"])
/ mean(agg_spl_all$dB[agg_spl_all$sound.files == i]) )


ggplot(agg_spl_all[agg_spl_all$sound.files %in% c("soundmeter", unique(agg_spl_all$sound.files)[2:6]), ], aes(x = orig.dB, y = adj_dB, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin=adj_dB - sd, ymax = adj_dB + sd, color = sound.files),width=.2, position=pd) +
    geom_line(position=pd) +
    scale_color_viridis_d() +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound decrease factor")


ggplot(agg_spl_all[agg_spl_all$sound.files %in% c("soundmeter", unique(agg_spl_all$sound.files)[7:10]), ], aes(x = orig.dB, y = adj_dB, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin=adj_dB - sd, ymax = adj_dB + sd, color = sound.files),width=.2, position=pd) +
    geom_line(position=pd) +
    scale_color_viridis_d() +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound decrease factor")

ggplot(agg_spl_all[agg_spl_all$sound.files %in% c("soundmeter", unique(agg_spl_all$sound.files)[11:16]), ], aes(x = orig.dB, y = adj_dB, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin=adj_dB - sd, ymax = adj_dB + sd, color = sound.files),width=.2, position=pd) +
    geom_line(position=pd) +
    scale_color_viridis_d() +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound decrease factor")
```


## Calibrating audiomoths using addition of mean amplitude difference compared to soundmeter

These plots show the SPL estimated by each audiomoths after calibration

```{r audiomoth calibration, eval = TRUE}
agg_spl <- read.csv(file.path(calibration.path, "spl_by_audiomoth_calibration_sounds.csv"))

names(agg_spl)[3] <- "dB"
agg_spl_sm$sound.files <- "soundmeter"
agg_spl_sm$label <- sort(as.numeric(unique(agg_spl$label)))

agg_spl_all <- rbind(agg_spl_sm, agg_spl)

agg_spl_all$orig.dB <- factor(agg_spl_all$label, levels = unique(agg_spl_all$label))

agg_spl_all$adj_dB <- agg_spl_all$dB

agg_spl_all <- agg_spl_all[agg_spl_all$label > 0.01, ]

# remove those with 0.1 not working
agg_spl_all <- agg_spl_all[!(agg_spl_all$label <= 0.1 & agg_spl_all$sound.files == "calibration_LM_A41A"), ]

agg_spl_all <- agg_spl_all[!(agg_spl_all$label < 0.04 & agg_spl_all$sound.files == "calibration_LM_AAE"), ]

agg_spl_all <- agg_spl_all[!(agg_spl_all$label < 0.04 & agg_spl_all$sound.files == "calibration_M_A41A"), ]


for(i in unique(agg_spl_all$sound.files[agg_spl_all$sound.files != "soundmeter"]))
  agg_spl_all$adj_dB[agg_spl_all$sound.files == i] <- agg_spl_all$adj_dB[agg_spl_all$sound.files == i] +
mean(agg_spl_all$dB[agg_spl_all$sound.files == "soundmeter" & agg_spl_all$label %in% agg_spl_all$label[agg_spl_all$sound.files == i]]) -
  mean(agg_spl_all$dB[agg_spl_all$sound.files == i])


ggplot(agg_spl_all[agg_spl_all$sound.files %in% c("soundmeter", "calibration_LM_A1A", "calibration_LM_A3C", "calibration_LM_A56", "calibration_LM_A61",  "calibration_LM_AAE"), ], aes(x = orig.dB, y = adj_dB, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin=adj_dB - sd, ymax = adj_dB + sd, color = sound.files),width=.2, position=pd) +
    geom_line(position=pd) +
    scale_color_viridis_d() +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound decrease factor")


ggplot(agg_spl_all[agg_spl_all$sound.files %in% c("soundmeter", "calibration_LM_ABB", "calibration_LM_AC8", "calibration_LM_AED", "calibration_LM_A41A"), ], aes(x = orig.dB, y = adj_dB, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin=adj_dB - sd, ymax = adj_dB + sd, color = sound.files),width=.2, position=pd) +
    geom_line(position=pd) +
    scale_color_viridis_d() +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound decrease factor")

ggplot(agg_spl_all[agg_spl_all$sound.files %in% c("soundmeter", "calibration_M_A1A", "calibration_M_A3C", "calibration_M_A41A", "calibration_M_A61", "calibration_M_AAE", "calibration_M_AC8"), ], aes(x = orig.dB, y = adj_dB, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin=adj_dB - sd, ymax = adj_dB + sd, color = sound.files),width=.2, position=pd) +
    geom_line(position=pd) +
    scale_color_viridis_d() +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound decrease factor")
```

```{r audiomoth calibration 3 for presentation, eval = TRUE}

agg_spl_all$am <- as.character(as.numeric(as.factor(agg_spl_all$sound.files)))
agg_spl_all$am[agg_spl_all$sound.files == "soundmeter"] <- "soundmeter"
agg_spl_all$am <- factor(agg_spl_all$am, levels = c("soundmeter", 1:16))
agg_spl_all$line.size <- ifelse(agg_spl_all$am == "soundmeter", 2, 1)

ggplot(agg_spl_all[agg_spl_all$orig.dB %in% c(0.0316, 0.1, 0.3162, 1), ], aes(x = orig.dB, y = adj_dB, group = am, color = am, size = line.size)) +
    geom_errorbar(aes(ymin=adj_dB - sd, ymax = adj_dB + sd, color = am),width=.2, position=pd) +
    geom_line(position=pd) +
    scale_color_viridis_d() +
    geom_point(position=pd, size=1) +
  theme_classic(base_size = 30) + labs(x = "Calibration sound SPL (dB)", y = "SPL measured (dB)", color = "") + guides(size = "none") +
    theme(text=element_text(size=30,  family="Calibri light"))


```


## Error margin after adjusting SPL from audiomoths
```{r error margin after adjusting SPL from audiomoths, eval = TRUE}
agg_spl_all <- agg_spl_all[order(agg_spl_all$sound.files, agg_spl_all$label), ]

mean_diff_dB_l <- lapply(unique(agg_spl_all$sound.files), function(x){  
  diffs <- agg_spl_all$dB[agg_spl_all$sound.files == "soundmeter" & agg_spl_all$label %in% agg_spl_all$label[agg_spl_all$sound.files == x]] - agg_spl_all$dB[agg_spl_all$sound.files == x] 
    
  out_df <- data.frame(sound.files = x, mean.diff = mean(diffs), sd.diff = sd(diffs))

  return(out_df)
})

mean_diff_dB <- do.call(rbind, mean_diff_dB_l)

mean_diff_dB$sd.diff[16] <- max(agg_spl_sm$sd)


mean_diff_dB <- mean_diff_dB[order(mean_diff_dB$sd.diff), ]

mean_diff_dB$sound.files <- factor(mean_diff_dB$sound.files, levels = mean_diff_dB$sound.files[order(mean_diff_dB$mean.diff)])

ggplot(mean_diff_dB, aes(x = sound.files, y = mean.diff, group = sound.files, color = sound.files)) +
    geom_errorbar(aes(ymin = mean.diff - sd.diff, ymax = mean.diff + sd.diff, color = sound.files),width=.2, position=pd) +
    geom_line(position=pd) +
    scale_color_viridis_d(end = 0.8) +
    geom_point(position=pd, size=1) +
  theme_classic() + labs(x = "Calibration sound decrease factor") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

# SPL adjustment for amplitude calibration 

```{r spl adjusting, eval = TRUE}

knitr::kable(mean_diff_dB)

write.csv(mean_diff_dB, "./output/spl_constant_for_audiomoth_calibration.csv", row.names = FALSE)

```


```{r test for detecting saturation, eval = FALSE}

saturated_df <- data.frame(sound.files = c("calibration_LM_A1A.WAV", "calibration_LM_A1A.WAV", "calibration_LM_A3C.WAV", "calibration_LM_A3C.WAV", "calibration_LM_A61.WAV", "calibration_LM_AED.WAV", "calibration_LM_AED.WAV", "calibration_M_A1A.WAV", "calibration_M_A1A.WAV", "calibration_M_A1A.WAV", "calibration_M_A1A.WAV", "calibration_M_A61.WAV", "calibration_M_A61.WAV", "calibration_M_A61.WAV", "calibration_M_A61.WAV", "calibration_M_AAE.WAV", "calibration_M_AAE.WAV", "calibration_M_AAE.WAV", "calibration_M_AAE.WAV", "calibration_M_AC8.WAV", "calibration_M_AC8.WAV", "calibration_M_AC8.WAV", "calibration_M_AC8.WAV", "calibration_M_AC8.WAV", "calibration_M_AC8.WAV", "calibration_M_AC8.WAV", "calibration_M_AC8.WAV"),
                           selec = c(432, 472, 2, 12, 2, 2, 32, 42, 82, 172, 352, 12, 52, 222, 302, 12, 82, 222, 262, 32, 62, 252, 282, 33, 63, 253, 283),
                           type = c("fine", "fine", "partial_saturation", "fine", "fine", "partial_saturation", "fine", "saturated", "saturated", "saturated",  "saturated",  "saturated",  "saturated", "saturated",  "saturated", "saturated", "saturated", "saturated", "saturated", "saturated", "saturated", "saturated", "saturated", "fine", "fine", "fine", "fine")) 

saturated_df$sound.files <- gsub("\\.WAV$", "", saturated_df$sound.files)
saturated_df$sf.selec <- paste(saturated_df$sound.files, saturated_df$selec, sep = "-")

alg.tests <- readRDS(file.path(calibration.path, "/sel_tab_audiomoth_calibration.RDS"))

alg.tests$sf.selec <- paste(alg.tests$sound.files, alg.tests$selec, sep = "-")

# extract examples for detecting saturation
sat_est <- alg.tests[alg.tests$sf.selec %in% saturated_df$sf.selec, ]

alg.tests$sf.selec <- saturated_df$sf.selec <- NULL

sat_est <- merge(sat_est, saturated_df, by = c("sound.files", "selec"))

sat_est <- fix_extended_selection_table(X = sat_est, Y = alg.tests)

sat <- detec_saturation(X = sat_est, parallel = 20)

sat[order(sat$type), ]

```

## Session information
```{r}

sessionInfo()

```

